\scshape Topics in mechanism design

### 2022-08-19 ch1.tex
Add files via upload

The above argument has a big hole: given our completely arbitrary mechanism $(q,p)$, there is no reason why the agent's reporting payoff $f(r,t) = t q(r) - p(r)$ should be a differentiable function of $r$, in which case the derivative
%
\begin{equation*}
	\left. \frac{\dd}{\dd m} f( t+m, t ) \right|_{m=0}
	\equiv f_1(t,t)
\end{equation*}
%
simply does not exist.
But the above argument \emph{can} be made rigorous: it really is true that the envelope formula is a fancy way of rewriting `local optimality' (suitably defined; see \textcite{Sinander2021}).
Alternatively and more traditionally,
one can use a completely different (unintuitive but elegant) argument to derive the envelope formula \parencite{MilgromSegal2002}.
Either way, we have learned:

@@

\paragraph{The literature.}
The modern envelope theorem is due to \textcite{MilgromSegal2002};
the more intuitive treatment here is based on \textcite{Sinander2021}.
The theorem applies to any parametrised optimisation problem in which an action $x$ is chosen from an arbitrary choice set $\mathcal{X}$ to maximise an objective $f(x,t)$,
where the only assumptions needed are that the parameter $t$ belongs to a connected space (without loss of generality, the unit interval $[0,1]$) and that $f(x,\cdot)$ is differentiable and satisfies a mild further condition (which rules out a badly-behaved derivative).%
	\footnote{\Cref{remark:env_powerful} illustrated the role of connectedness.
	If differentiability is dropped, we still obtain an inequality envelope formula involving directional (or Dini) derivatives; see e.g. \textcite{CarbajalEly2013}.}

@@

\paragraph{The literature.}
The Spence--Mirrlees lemma has been extended in various ways.
At the highest level of generality, the `outcome' $q$ belongs to an abstract space $Q$ ($=[0,1]$ in the text) equipped with a partial order,
and the agent's payoff is some function $f(q,p,t)$ ($=tq-p$ in the text).
Versions of the result go through whenever $f$ satisfies the \emph{Spence--Mirrlees (`single-crossing') condition,}
defined in terms of how different types' indifference curves in $q$--$p$ space cross each other.%
	\footnote{When preferences have the quasi-linear form $f(q,p,t) = g(q,t) - p$, Spence--Mirrlees requires precisely that $g$ be \emph{supermodular.}
	In general, a broader (ordinal) notion of supermodularity (or `complementarity') characterises the Spence--Mirrlees condition \parencite[][Theorem 3]{MilgromShannon1994}.}
See \textcite[§4]{Sinander2021} for an overview of such results (plus a new, general result).

>>>>

The above argument has a big hole: given our completely arbitrary mechanism $(q,p)$, there is no reason why the agent's reporting payoff $f(r,t) = t q(r) - p(r)$ should be a differentiable function of $r$, in which case the derivative
%
\begin{equation*}
	\left. \frac{\dd}{\dd m} f( t+m, t ) \right|_{m=0}
	\equiv f_1(t,t)
\end{equation*}
%
simply does not exist.
But the above argument \emph{can} be made rigorous: it really is true that the envelope formula is a fancy way of rewriting `local optimality' (suitably defined; see \textcite{Sinander2022}).
Alternatively and more traditionally,
one can use a completely different (unintuitive but elegant) argument to derive the envelope formula \parencite{MilgromSegal2002}.
Either way, we have learned:

@@

\paragraph{The literature.}
The modern envelope theorem is due to \textcite{MilgromSegal2002};
the more intuitive treatment here is based on \textcite{Sinander2022}.
The theorem applies to any parametrised optimisation problem in which an action $x$ is chosen from an arbitrary choice set $\mathcal{X}$ to maximise an objective $f(x,t)$,
where the only assumptions needed are that the parameter $t$ belongs to a connected space (without loss of generality, the unit interval $[0,1]$) and that $f(x,\cdot)$ is differentiable and satisfies a mild further condition (which rules out a badly-behaved derivative).%
	\footnote{\Cref{remark:env_powerful} illustrated the role of connectedness.
	If differentiability is dropped, we still obtain an inequality envelope formula involving directional (or Dini) derivatives; see e.g. \textcite{CarbajalEly2013}.}

@@

\paragraph{The literature.}
The Spence--Mirrlees lemma has been extended in various ways.
At the highest level of generality, the `outcome' $q$ belongs to an abstract space $Q$ ($=[0,1]$ in the text) equipped with a partial order,
and the agent's payoff is some function $f(q,p,t)$ ($=tq-p$ in the text).
Versions of the result go through whenever $f$ satisfies the \emph{Spence--Mirrlees (`single-crossing') condition,}
defined in terms of how different types' indifference curves in $q$--$p$ space cross each other.%
	\footnote{When preferences have the quasi-linear form $f(q,p,t) = g(q,t) - p$, Spence--Mirrlees requires precisely that $g$ be \emph{supermodular.}
	In general, a broader (ordinal) notion of supermodularity (or `complementarity') characterises the Spence--Mirrlees condition \parencite[][Theorem 3]{MilgromShannon1994}.}
See \textcite[§4]{Sinander2022} for an overview of such results (plus a new, general result).

### 2022-08-19 ch4.tex
Add files via upload

	As \textcite{GuoShmaya2021} point out,
	we get a sharper revelation principle if we assume instead that the agent can propose as many allocations as she likes:
	in that case, we may further limit our attention to those IC direct mechanisms that instruct the agent to propose \emph{all} of the allocations that she claimed (in her report) to be available.
	This is an instance of Bull and Watson's (\citeyear{BullWatson2007}) revelation principle for settings with hard evidence.}
% (We shall consider revelation principles for hard evidence in the next chapter.)
We shall therefore rely on a \emph{taxation principle} instead.

@@

\paragraph{The literature.} The model is due to \textcite{ArmstrongVickers2010}. The perspective here (which is quite different from that of the original paper) is one that I learned from Yingni Guo \parencite[see][]{GuoShmaya2021}.

@@

The art of variational calculus is to identify the `right' perturbations $d_\eps$.
This should be guided by economic intuition; the formalism should (and will) follow.
There are many candidates that might seem natural.
For example, we may add $\eps$ on a small interval $[u',u'+\delta)$:
%
\begin{equation*}
	d^{u',\delta}_\eps(u)
	\coloneqq
	\begin{cases}
		d(u) + \eps 
		&\text{for $u \in [u',u'+\delta)$} \\
		d(u)
		&\text{for $u \notin [u',u'+\delta)$.} 
	\end{cases}
\end{equation*}
%
By letting $\eps \to 0$, we get a first-order condition for each $u'$ and $\delta$. To obtain a more parsimonious family of first-order conditions, let $\delta$ vanish, too:
%
\begin{equation*}
	\left. \frac{\dd}{\dd \delta} \left. \frac{\dd}{\dd \eps}
	\Phi\left( d^{u',\delta}_\eps \right)
	\right|_{\eps=0} \right|_{\delta=0}
	= 0
	\quad \text{for each $u' \in \R_+$.}
\end{equation*}
%
This is a pretty common Euler equation.
Another natural (but rare) perturbation is $d_\eps(u) = d(u+\eps)$.
These are just two examples of perturbations;
both are useful,%
	\footnote{E.g. the former yields the Euler equation in \textcite{sfb}, and the latter constitutes the main definition in \textcite{Sinander2021}.}
and there are many others.

@@

\paragraph{The literature.}
\textcite{ArmstrongVickers2010} solve for the optimal $d$ using optimal control.
A note: they parametrise their model so that the principal's payoff is $v + \alpha u$ for some $\alpha \geq 0$, so their results are stated somewhat differently.
They derive comparative statics with respect to $\alpha$ (an `ally principle': more aligned preferences lead to more permissive delegation),
consider an extension in which allocations are generated by the agent via costly and unobservable effort (moral hazard),
and show that monetary transfers to the agent may not be used even if available.
They also provide an example of how the principal can do better if the agent is able to propose more than one allocation; this issue is studied further by \textcite{GuoShmaya2021}.

@@

\paragraph{The literature.}
\textcite{ArmstrongVickers2010} noted the benefits of randomisation, but did not study random mechanisms systematically.
\textcite{GuoShmaya2021} make progress by replacing our `Bayesian' objective (the principal has some fixed belief about how $A$ is distributed)
with an uncertainty-averse `minmax regret' objective.
This allows them to obtain a clean solution,
and (more importantly?) some nice qualitative insights about how the principal should optimally utilise her power to randomise.

>>>>

	As \textcite{GuoShmaya2022} point out,
	we get a sharper revelation principle if we assume instead that the agent can propose as many allocations as she likes:
	in that case, we may further limit our attention to those IC direct mechanisms that instruct the agent to propose \emph{all} of the allocations that she claimed (in her report) to be available.
	This is an instance of Bull and Watson's (\citeyear{BullWatson2007}) revelation principle for settings with hard evidence.}
% (We shall consider revelation principles for hard evidence in the next chapter.)
We shall therefore rely on a \emph{taxation principle} instead.

@@

\paragraph{The literature.} The model is due to \textcite{ArmstrongVickers2010}. The perspective here (which is quite different from that of the original paper) is one that I learned from Yingni Guo \parencite[see][]{GuoShmaya2022}.

@@

The art of variational calculus is to identify the `right' perturbations $d_\eps$.
This should be guided by economic intuition; the formalism should (and will) follow.
There are many candidates that might seem natural.
For example, we may add $\eps$ on a small interval $[u',u'+\delta)$:
%
\begin{equation*}
	d^{u',\delta}_\eps(u)
	\coloneqq
	\begin{cases}
		d(u) + \eps 
		&\text{for $u \in [u',u'+\delta)$} \\
		d(u)
		&\text{for $u \notin [u',u'+\delta)$.} 
	\end{cases}
\end{equation*}
%
By letting $\eps \to 0$, we get a first-order condition for each $u'$ and $\delta$. To obtain a more parsimonious family of first-order conditions, let $\delta$ vanish, too:
%
\begin{equation*}
	\left. \frac{\dd}{\dd \delta} \left. \frac{\dd}{\dd \eps}
	\Phi\left( d^{u',\delta}_\eps \right)
	\right|_{\eps=0} \right|_{\delta=0}
	= 0
	\quad \text{for each $u' \in \R_+$.}
\end{equation*}
%
This is a pretty common Euler equation.
Another natural (but rare) perturbation is $d_\eps(u) = d(u+\eps)$.
These are just two examples of perturbations;
both are useful,%
	\footnote{E.g. the former yields the Euler equation in \textcite{sfb}, and the latter constitutes the main definition in \textcite{Sinander2022}.}
and there are many others.

@@

\paragraph{The literature.}
\textcite{ArmstrongVickers2010} solve for the optimal $d$ using optimal control.
A note: they parametrise their model so that the principal's payoff is $v + \alpha u$ for some $\alpha \geq 0$, so their results are stated somewhat differently.
They derive comparative statics with respect to $\alpha$ (an `ally principle': more aligned preferences lead to more permissive delegation),
consider an extension in which allocations are generated by the agent via costly and unobservable effort (moral hazard),
and show that monetary transfers to the agent may not be used even if available.
They also provide an example of how the principal can do better if the agent is able to propose more than one allocation; this issue is studied further by \textcite{GuoShmaya2022}.

@@

\paragraph{The literature.}
\textcite{ArmstrongVickers2010} noted the benefits of randomisation, but did not study random mechanisms systematically.
\textcite{GuoShmaya2022} make progress by replacing our `Bayesian' objective (the principal has some fixed belief about how $A$ is distributed)
with an uncertainty-averse `minmax regret' objective.
This allows them to obtain a clean solution,
and (more importantly?) some nice qualitative insights about how the principal should optimally utilise her power to randomise.
